---
title: "Homework 1 Assignment"
author: flrnclee
output: html_document
published: false
tags: hw1
---

========================================================

# About the Data

The data set is a subset of the complete 2004 NYC HANES dataset. All survey participants are adults, ages 20+. Because information regarding eating behaviors, occupation, and BMI were all found in different survey portions, I used SAS to merge based on uniuqe SP_ID. SAS was also used to remove survey respondents with missing values on the variables of interest and export the data into a .CSV file. 

# Importing the Data

The dataset is located in a different folder. I will first retrieve the data. 
```{r}
setwd("C:/Users/Flo/Desktop/Data Viz 2014/Assignment1")
nychanes <- read.csv("nychanes.csv", stringsAsFactors=FALSE)
```

Explore the dataset.

```{r}
head(nychanes)
```

# Dataset Variables in nychanes 

(1) **"eatout"** is a composite score based on DBQ090G and DBQ090Q. It is a discrete numeric variable

* Numeric value corresponds to the number of times the survey respondent ate meals prepared in a restaurant (includes take-out orders)

(2) ***"occgrp"*** is a categorical variable based off of OCCREC and expressed numerically as a nominal numeric variable.

* occgrp=1 corresponds to managerial and professional specialty 
* occgrp=2 corresponds to sales
* occgrp=3 corresponds to technical and administrative 
* occgrp=4 corresponds to service
* occgrp=5 corresponds to farming, forestry, and fishing
* occgrp=6 corresponds to precision production, craft, and repair
* occgrp=7 corresponds to operators, fabricators, and laborers

(3) ***"BMXBMI"*** is a continuous variable that corresponds to the survey respondent's BMI based off of their height and weight. Another way of looking at BMI is through ***"BMI2"***, which is an ordinal categorical variable. 

* BMI2=1 corresponds to underweight
* BMI2=2 corresponds to normal weight
* BMI2=3 corresponds to overweight 
* BMI2=4 corresponds to obese

# Renaming Variables of Interest in nychanes

```{r}
colnames(nychanes)
newCol <- c("id", "eatout_bin", "eatout_num", "occ", "occgrp", "eatout", "bmi", "bmi_grp")
colnames(nychanes) <- newCol          
```

# Creating Initial Plot 

I am primarily interested to see whether there is some visual association between eating out and BMI. I expect to see higher frequency of eating out with higher BMI readings due to the high sodium, high fat content of most restaurant foods. 

(Later, I will add in occupational group to explore whether occupation affects this association.)

```{r}
colnames(nychanes)
newCol <- c("id", "eatout_bin", "eatout_num", "occ", "occgrp", "eatout", "bmi", "bmi_grp")
colnames(nychanes) <- newCol          
```





