Title: "Homework 3 Assignment"
author: Andy Tang, awt2118
output: html_document
published: false
tags: hw3
========================================================
I am using two waves (2006 and 2007) of the AsiaBarometer Survey to compare responses on how proud participants are of being their country's people.

Q18 in 2006 AsiaBarometer Survey and Q19 in 2007 AsiaBarometer Survey
How proud are you of being [YOUR COUNTRY’S PEOPLE]? (SA)
1 Very proud
2 Somewhat proud
3 Not really proud
4 Not proud at all
9 Don't know

```{r }
library(ggplot2)
library (plyr)
```



Read in the survey
```{r fig.width=7, fig.height=6}
readsurvey <- function(name){ location <- paste("/Users/awtang/Desktop/qmssviz/qmssviz",name, sep="/")
survey <- read.table (location, header=T)
return(survey)}

asia2006<-readsurvey("AsiaBarometer2006.txt")
asia2007<-readsurvey("AsiaBarometer2007.txt")

```

Omit NAs
```{r echo=FALSE}

na.omit(asia2006$Q18)
na.omit(asia2007$Q19)

```

```{r}
table (asia2006$Q18)
table (asia2007$Q19)
```



Histograms
```{r}
qplot(Q18, data = asia2006, fill = Country) + labs(title = "Proudness of Being Your Country's People in Asia 2006" ) + 
    labs(x = ("Proudness"), y = "Count")

qplot(Q19, data = asia2007, fill = Country) + labs(title = "Proudness of Being Your Country's People in Asia 2007") + 
    labs(x = ("Proudness"), y = "Count")

```

The countries in the 2007 wave of the survey have higher frequencies of respondents saying they are "very proud" of being their respective country's people.