Title: Ben Lee R-Tutorial Assignment
published: false
tags: hw2 sdfsdf
========================================================
Load in Data
```{r}
dat<-read.csv("C:/Users/blee/Documents/QMSS Fall 2014/DataViz/qmssviz/Lab1/lgas.csv",header = TRUE)
```



From the Nigeria Facilities Inventory data (`sample_data` from section 1.5.1),select all facilities located in the Southern zones of Nigeria.  

```{r}
fac <- read.csv("C:/Users/blee/Documents/QMSS Fall 2014/DataViz/qmssviz/Lab1/sample_health_facilities.csv", header=TRUE)
```

Incorporate the pop2006 column from the `lgas.csv` file into the new data.frame containing only those facilities located in Southern Nigeria. (Hint: your id column is `lga_id`). 

Choose zones in the South of Nigeria
```{r}
levels(fac$zone)
facsouth<-fac[fac$zone %in% c("South-South","Southeast","Southwest"),]
```
Create and abridged dataframe and merge the results
```{r}
pop2006dat<- dat[,c(1,3)] # Create abridged data frame
mergefacsouth<-merge(facsouth, pop2006dat, by = "lga_id") # Merge the results
```

Calculate the total number of full time nurses and doctors for all health facilities in each state.  
Create a data frame for total nurses by state
```{r}
nursestate<-aggregate(mergefacsouth$num_nurses_fulltime, by=list(mergefacsouth$state), sum, na.rm=TRUE)
colnames(nursestate)<-c("State","Nurses")
```

Create a data frame for total nurses by state
```{r}
docstate<-aggregate(mergefacsouth$num_doctors_fulltime, by=list(mergefacsouth$state), sum, na.rm=TRUE)
colnames(docstate)<-c("State","Doctors")
```
Merge the results
```{r}
nursedocstate<-merge(nursestate, docstate, by = "State") # Merge the results
```


Sort the resulting dataset by state population, in descending order.  

Append 2006 Population to previous dataframe
```{r}
totpop2006state<-aggregate(dat$pop_2006, by=list(dat$state), sum, na.rm=TRUE)
colnames(totpop2006state)<-c("State","Pop2006")
finaldat<-merge(nursedocstate, totpop2006state, by = "State") # Merge the results
```

Sort the final dataframe
```{r}
order(finaldat$Pop2006, decreasing=TRUE)
```

Show the results!  
```{r}
finaldat[order(finaldat$Pop2006, decreasing=TRUE),]





```