Title: "Homework 2 Assignment"
author: Andy Tang, awt2118
output: html_document
published: false
tags: hw2
========================================================

Reading in the files
```{r }
setwd ("/Users/awtang/Desktop/qmssviz/qmssviz/hw2")
lgas<-read.csv("lgas.csv")
nigeria<-read.csv ("sample_health_facilities.csv")
```

Produce a new data.frame, which has the following properties:From the Nigeria Facilities Inventory data (sample_data from section 1.5.1), select all facilities located in the Southern zones of Nigeria.
```{r fig.width=7, fig.height=6}
head(nigeria$zone)
south<-subset(nigeria, zone %in% c("South-South", "Southeast", "Southwest"))

```

Incorporate the pop2006 column from the lgas.csv file into the new data.frame containing only those facilities located in Southern Nigeria. 
```{r fig.width=7, fig.height=6}
lgaspop2006<-lgas[,c("lga_id","pop_2006")]
southpop2006<-merge (south, lgaspop2006, by="lga_id")
```

Calculate the total number of full time nurses and doctors for all health facilities in each state.
```{r}
doctors=aggregate(data=southpop2006, num_doctors_fulltime~state,FUN=sum)
nurses=aggregate(data=southpop2006, num_nurses_fulltime~state,FUN=sum)

```


Sort the resulting dataset by state population, in descending order.
```{r}
attach(southpop2006)
 orderedpop<-southpop2006[order(-pop_2006),]
detach(southpop2006)
```


Show results
```{r}
print(doctors)
print(nurses)
print(orderedpop)
```

