---
title: "Homework 2 assignment-JanetWenjuanDeng"
author: wd2234
output: html_document 
pulished: false   
tags: hw2 
---

1. Go to this link and download the file into the working directory.
2. Produce a new data.frame, which has the following properties:

Firstly, import sample data. 
```{r}
setwd("D:/Wenjuan Deng/Columbia/2014fall/data visualization/qmssgit/HW2")
file <- "sample_health_facilities.csv"
sample_data <- read.csv(file)
```
- From the Nigeria Facilities Inventory data (sample_data from section 1.5.1), select all facilities located in the Southern zones of Nigeria.
```{r}
southern<-subset(sample_data,zone %in% c("South-South","Southeast","Southwest"))
```

- Incorporate the pop2006 column from the lgas.csv file into the new data.frame containing only those facilities located in Southern Nigeria. 
```{r}
sample_data2 <- read.csv("lgas.csv")
sample_data3<-sample_data2[,c("lga_id","pop_2006")]
newdataframe<-merge(southern,sample_data3,by="lga_id")
````
Here we get a new data frame called newdataframe,and the first 6 records are as follows:
```{r, echo=FALSE}
head(newdataframe)
````

- Calculate the total number of full time nurses and doctors for all health facilities in each state.
```{r}
doctors<-by(newdataframe$num_doctors_fulltime,newdataframe$state,sum,na.rm=TRUE)
nurses<-by(newdataframe$num_nurses_fulltime,newdataframe$state,sum,na.rm=TRUE)
total<-doctors+nurses
````

```{r, echo=FALSE}
total
````

- Sort the resulting dataset by state population, in descending order.
```{r, echo=FALSE}
newdataframe[order(newdataframe$pop_2006, decreasing=TRUE), ]
```
