---
title: "R Tutorial Mini-Assignment"
author: "Elliot Cohen"
date: "September 25, 2014"
output: html_document
---
```{r RMD-global-options, include=FALSE}
## set R Markdown options. 
## Markdown defines the layout and format of the resulting document. 
## It is not R code, but allows you to hide/supress R code and output.
options(width=90)
# opts_chunk$set(warning=TRUE, message=FALSE , tidy=TRUE)
```

1. Go to this [link](http://bit.ly/1fj3sjD) and download the file into the working directory.  
2. Produce a new `data.frame`, which has the following properties:
    + From the Nigeria Facilities Inventory data (`sample_data` from section 1.5.1), select all facilities located in the Southern zones of Nigeria.
    + Incorporate the pop2006 column from the `lgas.csv` file into the new data.frame containing only those facilities located in Southern Nigeria. (Hint: your id column is `lga` or `lga_id`).  
3. Calculate the total number of full time nurses and doctors for all health facilities in each state.
    + Sort the resulting dataset by state population, in descending order.
    + Show the results!  

Read-in the raw data from .csv
```{r import-data}
lgas<-read.csv("labs/tutorial/lgas.csv")
str(lgas)
dim(lgas)

facilities<-read.csv("labs/tutorial/sample_health_facilities.csv")
str(facilities)
dim(facilities)
```

From the Nigeria Facilities Inventory data (`sample_health_facilities.csv` or 'sample_data' from section 1.5.1), select all facilities located in the Southern zones of Nigeria
```{r subset}
# What are the zone classifications for Nigeria?
levels(facilities$zone)
# grab facilities located in a Southern zone, e.g. a zone containing the regular expression "South". 
# This includes South-South, Southwest, and Southeast.
row<-grep("South", facilities$zone)
set<-facilities[row, ]
```

Incorporate the pop2006 column from the `lgas.csv` file into the new data.frame containing only those facilities located in Southern
```{r merge}
df<-merge(set, lgas[,3:4], by="lga")
```

Calculate the total number of full time nurses and doctors for all health facilities in each state.  
  - Sort the resulting dataset by state population, in descending order.  
  - Show the results! 
```{r stats}
library(plyr)
library(reshape2)
library(ggplot2)
states<-ddply(df, .(state), numcolwise(sum)) # summarize by state
states<-states[, -2] # remove the lga_id (not really a numeric value, and thus not additive). Should have been a factor or character string.
states[order(-states$pop_2006), ] # ordered table (by State population, decreasing)
```
Tufte always says that the best way to show a small data set with precision is with a table!