---
title: "Homework 3 - Kevin Gong"
output: html_document
date: "October 23, 2014"
published: false
tag: hw3
---

###**A**###

The survey I will be examining is the World Values Survey. Specifically, we'll be looking at the willingness of survey respondents to have homosexual neighbours in Wave 3 (1995-1998) and Wave 6 (2010-2014). The question appears on Card G in both waves and is labeled Question V60 in Wave 3 and V40 in Wave 6. There is a very slight different in working between the two waves, though we are not primarily concerned with the difference in wording.

**Wave 3:**
"On this list are various groups of people. Could you please sort out any that you would 
not like to have as neighbors? (CODE AN ANSWER FOR EACH) 
** Mentioned Not Mentioned
V51 People with a criminal record 1 2 
V52 People of a different race 1 2 
V53 Political extremists 1 2 
V54 Heavy drinkers 1 2 
V55 Emotionally unstable people 1 2 
V56 Muslims* 1 2 
V57 Immigrants/foreign workers 1 2 
V58 People who have AIDS 1 2 
V59 Drug addicts 1 2 
V60 Homosexuals 1 2"

**Wave 6:**
"On this list are various groups of people. Could you please mention any that you would not like to have as neighbors? (Code an answer for each group):

Mentioned Not mentioned
V36. Drug addicts 1 2
V37. People of a different race 1 2
V38. People who have AIDS 1 2
V39. Immigrants/foreign workers 1 2
V40. Homosexuals 1 2
V41. People of a different religion 1 2
V42. Heavy drinkers 1 2
V43. Unmarried couples living together 1 2
V44. People who speak a different language 1 2"

For these questions, note that a 1 indicates unwillingness to have as a neighbor. Also note that the groups of people listed across the two waves are different.



```{r}
load("~/qmssviz/hw3/WV3.RData")
WV3 = WV3_Data_rdata_v_2014_09_21
load("~/qmssviz/hw3/WV6.RData")
WV6 = WV6_Data_spss_v_2014_06_04
head(WV6)



```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

