---
title: "hw3-2014-10-23-sz2491"
author: "Siqian Zhu"
date: "2014-10-23"
output: html_document
published: false
tags: hw3
---

Decribe the question
V217. What is the highest educational level that you have attained? 
(use functional equivalent of the following, in given society; 
IF STUDENT, CODE HIGHEST LEVEL HE/SHE EXPECTS TO COMPLETE):
 1. No formal education 
 2. Incomplete primary school 
 3. Complete primary school 
 4. Incomplete secondary school: technical/vocational type 
 5. Complete secondary school: technical/vocational type 
 6. Incomplete secondary: university-preparatory type 
 7. Complete secondary: university-preparatory type 
 8. Some university-level education, without degree 
 9. University-level education, with degree 
 0. DK/NA 
 
set working directory and load packages

```{r}

setwd("/Users/ac/Documents/columbia/study/2014fall/datavisualization/github/qmssviz/hw3")

library(plyr)
library(reshape2)
library(ggplot2)
```

read data

```{r}
readdata <- function(){
load("~/Documents/columbia/study/2014fall/datavisualization/github/qmssviz/hw3/WV6_Data_rdata_v_2014_06_04.rdata")
wave6<-WV6_Data_spss_v_2014_06_04

load("~/Documents/columbia/study/2014fall/datavisualization/github/qmssviz/hw3/WV3_Data_rdata_v_2014_09_21.rdata")
wave3<-WV3_Data_rdata_v_2014_09_21
}

readdata()

```


sort data
choose countries in Asia, i.e. China, India, South Korea and Taiwan
```{r}
wave3a <- subset(wave3,select=c(V1,V2,V238,V217),V2==840|V2==156)
wave6a <- subset(wave6,select=c(V1,V2,V262,V248),V2==840|V2==156)


```


plot data

```{r}


```

trial

```{r}


```