---
title: "Nitkin HW1"
output: html_document
---

My first step was to bring my data into R. For this homework, I used data from the beginning-of-year and end-of-year staff satisfaction surveys at the network of charter schools where I work.

The survey was administered via surveymonkey, so I exported the data into a csv that
I could then import into R.

```
Fall_Survey <- "Fall_Survey.csv"
Spring_Survey <- "Spring_Survey.csv"
Fall <- read.csv(Fall_Survey)
Spring <- read.csv(Spring_Survey)
```

I then did some exploring of my new data frames.

```
dim(Fall)
dim(Spring)
colnames(Fall)
colnames(Spring)
```

I decided that I wanted to track the change in the average score for three questions:
"I feel valued as part of our community," "School leadership makes thoughtful decisions," and "I am proud to work for this organization."

At this point I tried to create objects for the variables that I cared about, but got an error message in R. I think the data was perhaps in the wrong format?

```
Fall_Value <- Fall$I.feel.valued.as.part.of.our.community
Spring_Value <- Spring$I.feel.valued.as.part.of.our.community
mean(Fall_Value)
plot(Fall_Value,Spring_Value)
```

I did some googling and found a website that recommended that I melt the data in Excel, then import it into R already formatted for graphing. So that's what I did. 

```{r}
Graphable_Data <- "../Survey_Data_To_Graph.csv"
Graphdata <- read.csv(Graphable_Data)
dim(Graphdata)
colnames(Graphdata)
```

Now that I had my data in a melted format, I tried to create a very simple clustered bar graph. I created a new object with the average score for each question, which I called "Score." Then I used that data to create my barplot.

```{r}
Score <- Graphdata$Score
```

I was able to plot this data just fine, but when I tried to use the {r, echo=FALSE} command to display it, I got an error message that said that "object 'Score' found found in Calls"

```
barplot(Score, main="Satisfaction Over Time", ylab = "Average Response Score", names.arg=c("I feel valued - Fall", "I feel valued - Spring", "Principal is thoughtful - Fall", "Principal is thoughtful - Spring", "I am proud - Fall", "I am proud - Spring"),las=2)
```

Then I took this graph and tried to turn it into a clustered bar graph by adding colors and a legend. The colors appeared, but the legend didn't.

```{r, echo=FALSE}
barplot(Score, main="Satisfaction Over Time", ylab = "Average Response Score", names.arg=c("I feel valued - Fall", "I feel valued - Spring", "Principal is thoughtful - Fall", "Principal is thoughtful - Spring", "I am proud - Fall", "I am proud - Spring"),las=2, col=c("darkblue","red"), beside=TRUE, cex.names=0.5)
```

And that's as far as I got!
