---
title: "Nitkin HW1"
output: html_document
---

My first step was to bring my data into R. For this homework, I used data from the beginning-of-year and end-of-year staff satisfaction surveys at the network of charter schools where I work.

The survey was administered via surveymonkey, so I exported the data into a csv that
I could then import into R.

```{r}
Fall_Survey <- "../Fall_Survey.csv"
Spring_Survey <- "../Spring_Survey.csv"
Fall <- read.csv(Fall_Survey)
Spring <- read.csv(Spring_Survey)
```

I then did some exploring of my new data frames.

```
dim(Fall)
dim(Spring)
colnames(Fall)
colnames(Spring)
```

I decided that I wanted to track the change in the average score for three questions:
"I feel valued as part of our community," "School leadership makes thoughtful decisions," and "I am proud to work for this organization."

At this point I created new values for the three variables that I cared about.

```{r}
Fall_Value <- Fall$I.feel.valued.as.part.of.our.community
Fall_Leadership <-Fall$School.leadership.makes.thoughtful.decisions
Fall_Pride <-Fall$I.am.proud.to.work.for.this.organization
Spring_Value <- Spring$I.feel.valued.as.part.of.our.community
Spring_Leadership <-Spring$School.leadership.makes.thoughtful.decisions
Spring_Pride <-Spring$I.am.proud.to.work.for.this.organization
```

Next I needed to convert this data from factors into numeric data.

```{r}
NumFallValue <- as.numeric(Fall_Value)
NumFallLead <- as.numeric(Fall_Leadership)
NumFallPride <- as.numeric(Fall_Pride)
NumSpringValue <- as.numeric(Spring_Value)
NumSpringLead <- as.numeric(Spring_Leadership)
NumSpringPride <- as.numeric(Spring_Pride)
```

Unfortunately, the as.numeric() function was converting blank responses into the number 1 and the response "Strongly Disagree (1)" into 11. So I wrote the below code to generate new objects with the correct values.

```{r}
NumFallValue[NumFallValue == 1] <- NA
NumFallValue[NumFallValue == 11] <- 1
NumFallValueFinal <- na.omit(NumFallValue)
```

I then repeated this for the other five variables.

```{r}
NumSpringValue[NumSpringValue == 1] <- NA
NumSpringValue[NumSpringValue == 11] <- 1
NumSpringValueFinal <- na.omit(NumSpringValue)
NumFallLead[NumFallLead == 1] <- NA
NumFallLead[NumFallLead == 11] <- 1
NumFallLeadFinal <- na.omit(NumFallLead)
NumSpringLead[NumSpringLead == 1] <- NA
NumSpringLead[NumSpringLead == 11] <- 1
NumSpringLeadFinal <- na.omit(NumSpringLead)
NumFallPride[NumFallPride == 1] <- NA
NumFallPride[NumFallPride == 11] <- 1
NumFallPrideFinal <- na.omit(NumFallPride)
NumSpringPride[NumSpringPride == 1] <- NA
NumSpringPride[NumSpringPride == 11] <- 1
NumSpringPrideFinal <- na.omit(NumSpringPride)
```

Next I created a new object called "Graphdata" that had only six values: the mean scores for the six questions that I cared about.

```{r}
Graphdata <- c(mean(NumFallValueFinal),mean(NumSpringLeadFinal),mean(NumFallPrideFinal),mean(NumSpringPrideFinal),mean(NumFallLeadFinal),mean(NumSpringLeadFinal))
```

Finally I graphed that object. Viola! I had a bar graph.

```{r, echo=FALSE}
barplot(Graphdata, main="Satisfaction Over Time", ylab = "Average Response Score", names.arg=c("I feel valued - Fall", "I feel valued - Spring", "Principal is thoughtful - Fall", "Principal is thoughtful - Spring", "I am proud - Fall", "I am proud - Spring"),las=2, col=c("darkblue","red"), beside=TRUE, cex.names=0.5, ylim=c(0,10))
```

It appears that staff satisfaction decreased over time on all three of the questions that I pulled out. Bad news for management :-(
