
title: "hw3_Lanxiang_Li_ll2937"
author: "Lanxiang Li"
date: "October 21, 2014"
output: html_document
published:false
tags:hw2

# A. Identify two surveys
I'm going to use two waves (wave2 and wave6)in the World Values Survey to examine the same survey question, simple but never out-dated:
Are you happy?

# B.Describe the graph(s)

# C.Write down the target shape
## year(wave):"1990-1994", "2010-2014"
## country: "United States","Sweden", "Spain"...
## survey answer: "very happy","rather happy","not very happy","not at all happy","no answer","don't know"
## proportion:

## D.

```{r}
setwd("~/Desktop/qmssviz/hw3")
load("/Users/carolineli/Desktop/qmssviz/hw3/WV2_Data_rdata_v_2014_04_28.rdata")
load("/Users/carolineli/Desktop/qmssviz/hw3/WV6_Data_rdata_v_2014_06_04.rdata")
dwv2 <- WV2_Data_rdata_v_2014_04_28
dwv6 <- WV6_Data_spss_v_2014_06_04
```

```{r}
str(dwv2$v18)
str(dwv6$V10)
dwv2$v18[dwv2$v18==1] <- "very happy"
dwv2$v18[dwv2$v18==2] <- "rather happy"
dwv2$v18[dwv2$v18==3] <- "not very happy"
dwv2$v18[dwv2$v18==4] <- "not at all happy"
dwv2$v18[dwv2$v18==-1] <- "don't know"
dwv2$v18[dwv2$v18==-2] <- "no answer"
dwv2$v18[dwv2$v18==-3] <- "not applicable"
dwv6$V10[dwv6$V10==1] <- "very happy"
dwv6$V10[dwv6$V10==2] <- "rather happy"
dwv6$V10[dwv6$V10==3] <- "not very happy"
dwv6$V10[dwv6$V10==4] <- "not at all happy"
dwv6$V10[dwv6$V10==-1] <- "don't know"
dwv6$V10[dwv6$V10==-2] <- "no answer"
dwv6$V10[dwv6$V10==-3] <- "not applicable"


```


