
title: "hw3_Lanxiang_Li_ll2937"
author: "Lanxiang Li"
date: "October 21, 2014"
output: html_document
published:false
tags:hw3

# A. Identify two surveys
I'm going to compare two waves (wave2 and wave6)in the World Values Survey in regards to the same survey question, simple but never out-dated:
Are you happy?

# B.Describe the graph(s)
I'm going to use stacked and grouped bar plots to compare the two waves of data.
Firstly, I'm going to use bar plots to compare different countries in the same wave.
Secondly, I'm going to use stacked bar plots to compare the two waves, horizontally and vertically.

# C.Write down the target shape
1.year(wave):"1990-1994", "2010-2014";
2.country: "Spain","Japan", "Mexico"...;
3.survey answer: "very happy","rather happy","not very happy","not at all happy","no answer","don't know";
4.proportion: the ratio of each level of happiness 

# D. Write a script

## import the data
```{r}
setwd("~/Desktop/qmssviz/hw3")
load("/Users/carolineli/Desktop/qmssviz/hw3/WV2_Data_rdata_v_2014_04_28.rdata")
load("/Users/carolineli/Desktop/qmssviz/hw3/WV6_Data_rdata_v_2014_06_04.rdata")
dwv2 <- WV2_Data_rdata_v_2014_04_28
dwv6 <- WV6_Data_spss_v_2014_06_04
```

## check the data
```{r}
dim(dwv2)
dim(dwv6)
str(dwv2$v18)
str(dwv6$V10)
```

## Rename the Level of Happiness
```{r}
dwv2$v18[dwv2$v18==1] <- "very happy"
dwv2$v18[dwv2$v18==2] <- "rather happy"
dwv2$v18[dwv2$v18==3] <- "not very happy"
dwv2$v18[dwv2$v18==4] <- "not at all happy"
dwv2$v18[dwv2$v18==-1] <- "don't know"
dwv2$v18[dwv2$v18==-2] <- "no answer"
dwv2$v18[dwv2$v18==-3] <- "not applicable"
dwv6$V10[dwv6$V10==1] <- "very happy"
dwv6$V10[dwv6$V10==2] <- "rather happy"
dwv6$V10[dwv6$V10==3] <- "not very happy"
dwv6$V10[dwv6$V10==4] <- "not at all happy"
dwv6$V10[dwv6$V10==-1] <- "don't know"
dwv6$V10[dwv6$V10==-2] <- "no answer"
dwv6$V10[dwv6$V10==-3] <- "not applicable"
table(dwv6$V10)
table(dwv2$v18)
```

## selecting data from 6 countries
```{r}
dwv2_new <- subset(dwv2,select=c("v1","v18","v2"),v2==156|v2==392|v2==410|v2==484|v2==566|v2==724)
dwv6_new <- subset(dwv6,select=c("V1","V10","V2"),V2==156|V2==392|V2==410|V2==484|V2==566|V2==724)
```

## Rename the countries
```{r}
dwv2_new$v2[dwv2_new$v2==156] <- "China"
dwv2_new$v2[dwv2_new$v2==392] <- "Japan"
dwv2_new$v2[dwv2_new$v2==410] <- "South Korea"
dwv2_new$v2[dwv2_new$v2==484] <- "Mexico"
dwv2_new$v2[dwv2_new$v2==566] <- "Nigeria"
dwv2_new$v2[dwv2_new$v2==724] <- "Spain"
dwv6_new$V2[dwv6_new$V2==156] <- "China"
dwv6_new$V2[dwv6_new$V2==392] <- "Japan"
dwv6_new$V2[dwv6_new$V2==410] <- "South Korea"
dwv6_new$V2[dwv6_new$V2==484] <- "Mexico"
dwv6_new$V2[dwv6_new$V2==566] <- "Nigeria"
dwv6_new$V2[dwv6_new$V2==724] <- "Spain"
```

## rename the columns
```{r}
colnames(dwv2_new) <-c("year","happiness","country")
colnames(dwv6_new) <-c("year","happiness","country")
dwv2_new$year <-"1992"
dwv6_new$year <-"2012"
```


## combine the two waves
```{r}
combine<-rbind(dwv2_new, dwv6_new)
dim(combine)
```

#E.Make it actually make your graph.
```{r}
library(ggplot2)
ggplot(combine, aes(x=year, fill=happiness)) + geom_bar(position="fill", binwidth=40) + facet_wrap(~country) + xlab("year") + ylab("happiness") + ggtitle("Comparison of happiness in different countries in 1992 and 2012") + scale_fill_discrete(name="proportion of\nhappiness")
```
## Summary:
From the grah, we can see that in some countries, there are more people feeling very happy(such as Mexico, Nigeria and South Korea), while in some countries there's a decrease. We can get similar conclusions such as differences in two waves in different countries from the grah. 

# F&G. write funciton calls
## lookatdata <- function(data=d)
{
head(d)
dim(d)
view(d)
}
## 

