---
title: "Dataviz Assignment 3: Data analysis and visualization"
author: "Ben Chow"
date: "22 October, 2014"
output: html_document
---

### **E, G** ###

**1. Clustering Analysis**

```{r, message = FALSE}
require("mice")
require("RColorBrewer")
WVS <- read.csv("WVS456_rel.csv")
# This next step should be exercised with caution: due to multiple joins and overwrites my csv file grew extra columns 
# If you ran my code with the original files on your own, it may not be necessary
# If you used the csv file provided in the Git folder then it will work.
WVS <- WVS[,-1:-2]

# Extract complete cases
wvs <- WVS[,1:6]
w <- WVS[complete.cases(wvs), ]
rm(WVS, wvs)

# k-means clustering
set.seed(100)
kclust4 <- kmeans(w[,1:6], 4)
kclust5 <- kmeans(w[1:6], 5)
```

**2. Tidying up of Results**

```{r}
# Vectors of religious measures
relig <- c("prvblf", "othblf", "pubact", "prvact", "iden1", "iden2")

# Results from analysis
w$kclust4 <- kclust4$cluster
w$kclust5 <- kclust5$cluster

# Tabulating results
tab41 <- aggregate(cbind(w$prvblf, w$othblf, w$pubact, w$prvact, w$iden1, w$iden2) ~ w$kclust4, FUN = mean, na.action=na.omit)
tab41 <- data.frame(tab41)

tab51 <- aggregate(cbind(w$prvblf, w$othblf, w$pubact, w$prvact, w$iden1, w$iden2) ~ w$kclust5, FUN = mean, na.action=na.omit)
tab51 <- data.frame(tab51)

# Some labels
colnames(tab41) <- colnames(tab51) <- c("clust", relig)

# Normalization of Scores across religious variables
for (i in 2:7) {
    frac <- c(9, 9, 6, 1, 2, 3)
    tab41[,i] <- tab41[,i]/frac[i-1]
    tab51[,i] <- tab51[,i]/frac[i-1]
}
```

**3. Reordering the Table**
```{r}
# Want to reorder results in decreasing order across columns
tab41 <- tab41[,c(2:7, 1)]
ord41 <- order(apply(tab41[1:6], 2, FUN = sum), decreasing = TRUE)

tab51 <- tab51[,c(2:7, 1)]
ord51 <- order(apply(tab51[1:6], 2, FUN = sum), decreasing = TRUE)

tab41 <- tab41[,c(ord41, 7)]
tab51 <- tab51[,c(ord51, 7)]

# Extracting information about denomination and appending it to the tables
tab42 <- aggregate(w$denom ~ w$kclust4, FUN = table, na.action=na.omit)
tab42 <- data.frame(tab42[,"w$denom"])
tab52 <- aggregate(w$denom ~ w$kclust5, FUN = table, na.action=na.omit)
tab52 <- data.frame(tab52[,"w$denom"])

x5 <- cbind(tab51, tab52)
ord52 <- order(apply(x5[1:6], 1, FUN = sum), decreasing = TRUE)
x5 <- x5[ord52,]

x4 <- cbind(tab41, tab42)
ord42 <- order(apply(x4[1:6], 1, FUN = sum), decreasing = TRUE)
x4 <- x4[ord42,]


```

**4. Preparation for Visualization**

```{r}
x5$clust <- factor(x5$clust, levels = ord52)
x4$clust <- factor(x4$clust, levels = ord42)

# Extraction of colors from IWantHue and RColorBrewer
color1 <- c("#2BA4D1", "#C0B00D", "#DC2975", "#29E887", "#FE93FE", "#7165A8")
color2 <- brewer.pal(5, "Set1")

# Labels to be inserted into the graphs (reordered)
lab <- c("God", "taxes", "attend", "prayer", "self", "general")
lab <- lab[ord51]
```

Below I display again the questions with their associated labels:

- **God**: How important is God in your life?
- **taxes**: How often is cheating on taxes justified, if you have the chance?
- **attend*: Apart from weddings and funerals, about how often do you attend religious services these days?
- **prayer**: Do you have moments of prayer and meditation from time to time?
- **self**: Independently of whether you attend religious services, how religious would you say you are?
- **general**: How important is religion in your life?

I also produce the corresponding table with a breakdown of the number of respondents from each of the religions (there were too few Muslims, Buddhists etc. so they were all lumped into "Other"). Note that the cluster numbers in the graph do not correspond to the ones in the first column, which are 'random' products of the k-means algorithm. So cluster 1 (as in the graph) corresponds with the first row, and so on. If I had more time I would have augmented the graph with this information as well, such as plotting a mini clustered bar chart underneath the points for Plot 1. 

```{r, echo = FALSE}
clusterno <- c(1:5)
cbind(clusterno, x5[8:11])
```


**5. Plot 1**

```{r, echo = FALSE}
plot(x5$clust, x5$Other, 
     type = "n", 
     frame.plot = F, 
     main = "Characteristics of different types of\n religious people", 
     ylim = c(0, 1), 
     xlab = "Cluster No.", ylab = "Score", 
     axes = F)
for(i in 1:5){
    axis(1, at = c(i), labels = c(i), lwd = 2, col.axis = color2[i], col.ticks = color2[i])    
}
axis(2, at = c(seq(0, 1, by = 0.2)), labels = seq(0, 1, by = 0.2), lwd = 0, las = 1)
for(i in 1:6){
    points(x5$clust, x5[,relig[i]], pch = 19, col = color1[i])
    lines(x5$clust, x5[,relig[i]], lwd = 2, col = color1[i], lty = 3)
}
legend(1, 0.45, legend = lab, col = color1, pch = 16, text.col = color1, box.lty = 0)
```

**6. Plot 2**

```{r, echo = FALSE}
plot(c(1:6), c(rep(5, 6)), 
     type = "n", 
     frame.plot = F, 
     main = "Religious scores by category", 
     xlim = c(1, 6), ylim = c(0, 1), 
     xlab = NA, ylab = "Score",
     axes = F)
for(i in 1:6){
    axis(1, at = c(i), labels = lab[i], lwd = 2, col.ticks = color1[i], col.axis = color1[i])
}
axis(2, at = c(seq(0, 1, by = 0.2)), labels = seq(0, 1, by = 0.2), lwd = 0, las = 1)
for(i in 1:5){
    lines(c(1:6), x5[i, 1:6], lwd = 2, col = color2[i])
    points(c(1:6), x5[i, 1:6], pch = 16, col = color2[i])    
}
legend(1, 0.39, legend = c("cluster 1", "cluster 2", "cluster 3", "cluster 4", "cluster 5"), col = color2, lty = 1, lwd = 2, text.col = color2, box.lty = 0)
```


** 7. Analysis **

From these graph it appears that we can begin to characterize the different types of religious people:

* Cluster 1: Can we call them the "truly" religious people? They score high on all counts, and would not embezzle money if they had the chance. Clearly the kind of people we need running for office. However, only less than a sixth (504 out of 3199) of all religious people belong to this category 
* Cluster 2: The majority. About two-fifths of the respondents fall into this category. These guys seem to be pretty grounded in religion, but will occasionally sleep in on Sundays and forget to do their daily quiet time. They seem very assured in their beliefs, but these manifest themselves slightly less strongly in their actions.
* Cluster 3: The bankers and financiers. Taxes susceptible to evasion only have their designers to blame. It's a free-for-all society. While cheating on taxes is fine for these guys, they still go to the church/synagogue/mosque/temple fairly often, and pray somewhat less than the majority. About one tenth of the people fall here.
* Cluster 4: Like the majority, except with very little free time. These guys probably work 80-hour-weeks and fast-paced lifestyles, which explains why they hardly have time to pray. Despite their lack of physical engagement, they somehow manage to maintain their religious beliefs rather strongly. Actually, wait. Maybe these guys are the real bankers, who work so hard and earn so much money they don't see a need to renege on their taxes.
* Cluster 5: The mess. Dear, dear, what have we here? Just over a tenth of respondents fall into the category of shame. Beliefs (private, public and general) all stayed high for the previous 4 categories, and here they all plummet. It appears that people who falter in their beliefs also correspondingly lack in action. Possibly those on the verge of apostasy? Or simply disaffected new-generation youths who have been brought up in a religious environment with no real spiritual engagement?

These visualizations are not meant to draw any firm conclusions about the reasons underlying these distributions. Rather, they are meant to provoke thought about what defines and characterizes religious adherence. So many people loosely brand themselves as being religious, but how many actually follow the Bible, Koran or Talmud? More importantly, if you yourself consider yourself religious, where would you fall?

The graphs open up a plethora of directions for further exploration. These categorizations would serve as an important control for analyses on the relationship between religion and other variables. One can think of many such examples on the fly, such as homosexual attitudes, political views and charitable giving.
