---
title: "Homework 2"
author: "Ryan Quan"
date: "September 25, 2014"
output: html_document
---
### Setup

```{r, message=FALSE}
setwd("~/mailman/qmssviz/hw2/_posts/")
library(plyr)
library(dplyr)
```

### Loading the Data

```{r}
lgas <- read.csv("lgas.csv")
niger <- read.csv("sample_health_facilities.csv")
```

### Column Names

```{r}
names(lgas)
```

```{r}
names(niger)
```

### Subsetting

Let's select facilities located in the Southern zones of Nigeria. We use the `unique` function to look at the different labels for zones and subset for the ones in the with the term "south".

```{r}
unique(niger$zone)
nigerSouth <- niger[niger$zone %in% 
                      c("South-South", "Southeast", "Southwest"), ]
```

We are interested in getting the `pop2006` column from the lgas dataset. Let's subset the data on the columns `pop_2006` and `lga_id` which we will need to join the datasets.

```{r}
pop2006 <- lgas[, c("lga_id","pop_2006")]
```

### Merging the Datasets

Here we use the `join` function from the `plyr` package to join the two datasets on `lga_id`.

```{r}
lgasSouth <- join(nigerSouth, pop2006, by = "lga_id")
```

### Summary

Let's calculate the total number of full time nurses and doctors for all health facilities in each state, sorted by the state population. We use the `dplyr` package for readability.

```{r}
lgasSouth %>%
    group_by(state) %>%
    summarise(num_nurses = sum(num_nurses_fulltime),
              num_doctors = sum(num_doctors_fulltime),
              population = sum(pop_2006)) %>%
    arrange(desc(population))
```


```