---
title: "Homework 2 Assignment"
author: "yw2668"
date: "October 13, 2014"
output: html_document
---

####1. Produce a new dataframe with the following properties:
####   Select all health facilities located in the Southern zones of Nigeria.
####   Incorporate LGA metadata into this new dataframe.

```{r}
setwd("/Users/Emily/Documents/Columbia/Data Visualization")
d_lgas<-read.csv("lgas.csv")
data<-read.csv("Health_Mopup_and_Baseline_NMIS_Facility.csv")
table(d_lgas$zone)
lgas_south<-d_south1<-subset(d_lgas, zone %in% c("South-South", "Southeast", "Southwest"))
newdata<-merge(lgas_south, data, by="unique_lga", all.x=TRUE)

```

####2. Calculate the total number of full time nurses and doctors in all health facilities, by state. Compute per capita and per square area values, as well.
####   Sort the first resultant dataset by state population, in descending order.
####   Show the results!

```{r}
library(plyr)
doctors<-ddply(newdata , .(state), summarise, sum=sum(num_doctors_fulltime,na.rm=TRUE))
nurses<-ddply(newdata , .(state), summarise, sum=sum(num_nurses_fulltime,na.rm=TRUE))
pop_2006<-ddply(newdata , .(state), summarise, sum=sum(pop_2006,na.rm=TRUE))
area<-ddply(newdata , .(state), summarise, sum=sum(area_sq_km,na.rm=TRUE))

d1<-merge(doctors, nurses,  by="state")
d2<-merge(d1, pop_2006,  by="state")
d3<-merge(d2,area, by="state")
names(d3)[2]<-"doctors"
names(d3)[3]<-"nurses"
names(d3)[4]<-"pop_2006"
names(d3)[5]<-"area"
d4<-d3[order(-d3$pop_2006), ]
print(d4)
result<-cbind(d3$state, d3[,2:3]/(d3$pop_2006))
names(result)[1]<-"state"
print(result)
```

