---
title: "Short R Assignment"
output: html_document
date: "September 24, 2014"
published: false
tag: hw1
---



Loading the data:

```{r,echo=FALSE}
lgas <- read.csv("~/qmssviz/labs/lgas.csv")
facilities <- read.csv("~/qmssviz/labs/tutorial/sample_health_facilities.csv")
head(lgas)
str(facilities)
head(facilities)
unique(facilities$zone)
```




From the Nigeria Facilities Inventory data, select all facilities located in the Southern zones of Nigeria.

```{r,echo=FALSE}
condition1 = which(facilities$zone==c("Southeast") | facilities$zone==c("South-South") | facilities$zone==c("Southwest"))
fac2 <- facilities[condition1,]
fac2

unique(lgas$zone)[2]
```




Incorporate the pop2006 column from the lgas.csv file into the new data.frame containing only those facilities located in Southern Nigeria. We used lga_id in both datasets to complete the merge.


```{r,echo=FALSE}
fac3 <- merge(fac2, lgas, by.x="lga_id", by.y="lga_id")
fac3
```



Calculate the total number of full time nurses and doctors for all health facilities in each state.

```{r,echo=FALSE}
unique(fac3$state.y)
health_counts = data.frame(unique(fac3$state.y))
health_counts[,2]=NA
#health_counts


for(i in 1:11) {
  
  tally = 0
  temp = as.vector(NULL)
  
  temp = which(as.character(fac3$state.y)==as.character(health_counts[i,1]))
  tally = sum(sum(fac3$num_nurses_fulltime[temp],na.rm=TRUE) + sum(fac3$num_doctors_fulltime[temp],na.rm=TRUE))
  
  
  health_counts[i,2] = tally
  
  
}

names(health_counts)[1] <- paste("State")
names(health_counts)[2] <- paste("Total Doctors + Nurses")
health_counts


```




Sort the resulting dataset by state population, in descending order.

```{r,echo=FALSE}
#fac4 <- fac3[order(-fac3$pop_2006),]
#fac4

unique(fac3$state.y)
statepop = data.frame(unique(fac3$state.y))
statepop[,2]=NA
#health_counts


for(i in 1:11) {
  
  tally = 0
  temp = as.vector(NULL)
  
  temp = which(as.character(fac3$state.y)==as.character(statepop[i,1]))
  tally = sum(sum(fac3$pop_2006[temp],na.rm=TRUE))
  
  
  statepop[i,2] = tally
  
  
}

names(statepop)[1] <- paste("State")
names(statepop)[2] <- paste("State Population")
statepop


finalset1 <- merge(health_counts, statepop, by.x="State", by.y="State")
#finalset1

finalset2 <- finalset1[order(-finalset1[,3]),]
finalset2
```





