---
title: "HW2"
author: "Liangquan Zhou lz2377"
date: "Wednesday, November 19, 2014"
output: html_document
---


## Now let's finish today's hw(2014-09-18). First load sample data.

```{r, cache=TRUE}
setwd("C:/Liangquan Zhou/Study/2014 fall/data visualization/hw1")
sample_data= read.csv("sample_health_facilities.csv") # read the .csv file
```

1. Select all facilities located in the southern zones of Nigeria.

```{r, cache=TRUE}
new_data= subset(sample_data, zone %in% c("Southwest","Southeast","South-South"))
```

2. Incorporate the pop2006 column from the `lgas.csv` file into the new data.frame containing only those facilities located in Southern Nigeria. (Hint: your id column is `lga_id`)

```{r, cache=TRUE}
lgas= read.csv("lgas.csv", stringsAsFactors=T)
new_data= merge(new_data, lgas[c("lga_id","pop_2006")],by= "lga_id")
```

And we can use `str` to see the new dataset:
```{r}
str(new_data)
```

3. Calculate the total number of full time nurses and doctors for all health facilities in each state.
```{r, cache=TRUE}
tapply(new_data$num_doctors_fulltime,  new_data$state,sum)
tapply(new_data$num_nurses_fulltime,  new_data$state,sum)
```

4. Sort the resulting dataset by state population, in descending order.
```{r,cache=TRUE}
data1=subset(new_data,select=c(num_doctors_fulltime,num_nurses_fulltime,pop_2006,state))
data1$state=as.factor(as.character(data1$state))
result=data.frame(tapply(data1$num_doctors_fulltime, data1$state,sum),
  tapply(data1$num_nurses_fulltime,  data1$state,sum),
  tapply(data1$pop_2006,  data1$state,sum))
names(result)=c("num_doctors_fulltime","num_nurses_fulltime","pop_2006")
result=result[order(result$pop_2006),]
```

The result is:
```{r,cache=TRUE}
result
```