---
title: "ECREEE_R_Training_1_Assignment"
author: "Yue Long"
date: "September 19, 2014"
output: html_document
---
First we recall the package plyr and read data from sample_health_facilities:
```{r}
library(plyr)
sample_health <- read.csv("~/Documents/qmssviz/Lab1/sample_health_facilities.csv")
```
Next we take out the data for places in the Southern areas:
```{r}
south <-subset(sample_health, zone%in%c("South-South", "Southeast", "Southwest"))
```
Then we read data from lgas and merge the population column in 2006 to the data before:
```{r}
lgas <- read.csv("~/Documents/qmssviz/lgas.csv", header=FALSE)
south_match <- subset(lgas, V1%in%south$lga_id)
pop_2006 <-data.frame(south_match$V1, south_match$V3)
colnames(pop_2006) <- "lga_id"
colnames(pop_2006)[2] <- "pop_2006" 
new <- merge(south, pop_2006, by="lga_id")
```

Since we are interested in summing up the number of fulltime nurses and doctors by state and sort the states using population, we may want to simplify things a little bit by taking out onlyl the four columns with information about states, nurses, doctors and population. 
```{r}
facilities <- data.frame(new$state, new$num_nurses_fulltime, new$num_doctors_fulltime, new$pop_2006)
```

Next we create a column for each different state and put the number of nurses, doctors and population in the vector. I want to know a different and easier way to this instead of inputing values for each single state.

```{r}
Abia<-subset(facilities, facilities$new.state=="Abia")
Abia<-c(sum(Abia$new.num_nurses_fulltime,na.rm=TRUE), sum(Abia$new.num_doctors_fulltime,na.rm=TRUE), sum(as.numeric(as.character(Abia$new.pop_2006)),na.rm=TRUE))
Abia_name <- "Abia"

Anambra<-subset(facilities, facilities$new.state=="Anambra")
Anambra<-c(sum(Anambra$new.num_nurses_fulltime,na.rm=TRUE), sum(Anambra$new.num_doctors_fulltime,na.rm=TRUE), sum(as.numeric(as.character(Anambra$new.pop_2006)),na.rm=TRUE))
Anambra_name <- "Anambra"

Cross_River<-subset(facilities, facilities$new.state=="Cross River")
Cross_River<-c(sum(Cross_River$new.num_nurses_fulltime,na.rm=TRUE), sum(Cross_River$new.num_doctors_fulltime,na.rm=TRUE), sum(as.numeric(as.character(Cross_River$new.pop_2006)),na.rm=TRUE))
Cross_River_name <- "Cross_River"

Delta<-subset(facilities, facilities$new.state=="Delta")
Delta<-c(sum(Delta$new.num_nurses_fulltime,na.rm=TRUE), sum(Delta$new.num_doctors_fulltime,na.rm=TRUE), sum(as.numeric(as.character(Delta$new.pop_2006)),na.rm=TRUE))
Delta_name <-"Delta" 

Edo<-subset(facilities, facilities$new.state=="Edo")
Edo<-c(sum(Edo$new.num_nurses_fulltime,na.rm=TRUE), sum(Edo$new.num_doctors_fulltime,na.rm=TRUE), sum(as.numeric(as.character(Edo$new.pop_2006)),na.rm=TRUE))
Edo_name <- "Edo" 

Ekiti<-subset(facilities, facilities$new.state=="Ekiti")
Ekiti<-c(sum(Ekiti$new.num_nurses_fulltime,na.rm=TRUE), sum(Ekiti$new.num_doctors_fulltime,na.rm=TRUE), sum(as.numeric(as.character(Ekiti$new.pop_2006)),na.rm=TRUE))
Ekiti_name <- "Ekiti"

Imo<-subset(facilities, facilities$new.state=="Imo")
Imo<-c(sum(Imo$new.num_nurses_fulltime,na.rm=TRUE), sum(Imo$new.num_doctors_fulltime,na.rm=TRUE), sum(as.numeric(as.character(Imo$new.pop_2006)),na.rm=TRUE))
Imo_name <- "Imo"

Lagos<-subset(facilities, facilities$new.state=="Lagos")
Lagos<-c(sum(Lagos$new.num_nurses_fulltime,na.rm=TRUE), sum(Lagos$new.num_doctors_fulltime,na.rm=TRUE), sum(as.numeric(as.character(Lagos$new.pop_2006)),na.rm=TRUE))
Lagos_name <- "Lagos"

Ogun<-subset(facilities, facilities$new.state=="Ogun")
Ogun<-c(sum(Ogun$new.num_nurses_fulltime,na.rm=TRUE), sum(Ogun$new.num_doctors_fulltime,na.rm=TRUE), sum(as.numeric(as.character(Ogun$new.pop_2006)),na.rm=TRUE))
Ogun_name <- "Ogun"

Osun<-subset(facilities, facilities$new.state=="Osun")
Osun<-c(sum(Osun$new.num_nurses_fulltime,na.rm=TRUE), sum(Osun$new.num_doctors_fulltime,na.rm=TRUE), sum(as.numeric(as.character(Osun$new.pop_2006)),na.rm=TRUE))
Osun_name <- "Osun"

Rivers<-subset(facilities, facilities$new.state=="Rivers")
Rivers<-c(sum(Rivers$new.num_nurses_fulltime,na.rm=TRUE), sum(Rivers$new.num_doctors_fulltime,na.rm=TRUE), sum(as.numeric(as.character(Rivers$new.pop_2006)),na.rm=TRUE))
Rivers_name <- "Rivers"
```

In the end we combine these vectors into a dataframe and sort them by population:
```{r}
summary <- rbind(Abia, Anambra, Cross_River, Delta, Edo, Ekiti, Imo, Lagos, Osun, Ogun, Rivers)
colnames(summary)[3] <- "pop_2006"
final <- summary[order(-summary[ ,3]),] 
final
```


