---
title: "140918 Data Visualization Assignment"
author: "Ben Chow"
date: "23 September, 2014"
output: html_document
---

In this assignment we were asked to present a dataframe obtained by manipulating data found in two csv files.

### Task 1 ###
+ **Download file into working directory**

After saving the relevant csv files into my working directory I execute:
```{r}
d <- read.csv("sample_health_facilities.csv")
```

### Task 2 ###
+ **from `sample_data` select facilities located in Southern zones**

The `grep` command searches the dataframe `d` and saves the corresponding entries. I build a new dataframe with only the southern facilities which i call `df`.

```{r}
south <- grep("South", d[,"zone"], fixed = TRUE)
df <- d[south, ]
```

### Task 3 ###
+ **Incorporate pop2006 column from other dataset into the dataframe**
```{r}
# Load data
pop <- read.csv("lgas.csv")
# Extract population values and merge
pop06 <- pop[df$lga_id, c("lga_id", "pop_2006")]
df <- merge(df, pop06, by = "lga_id")
```

### Task 4 ###
+ **Calculate total number of full time nurses and doctors in each state**

For the next task and others we will need the package `"plyr"` for the commands `ddply` and `join`. Here I save the totals in a dataframe `totmed`.

```{r, message=FALSE}
require("plyr")
totmed <- ddply(df, "state", summarise, Doctors = sum(num_doctors_fulltime, na.rm = TRUE), Nurses = sum(num_nurses_fulltime, na.rm = TRUE))
```

### Task 5 ###
+ **Sort the resulting dataset by state population, in decreasing order**
```{r}
# Calculating State Populations
stpop <- aggregate(pop$pop_2006 ~ pop$state, FUN = sum)
colnames(stpop) <- c("state", "State Pop")

# Merging data
totmed <- join(totmed, stpop, by = "state", type = "left", match = "first")

# Sort by state population
totmed.sort <- totmed[order(totmed[, 4], decreasing = TRUE), ]
```


### Final Output ###
```{r, collapse=TRUE}
totmed.sort
```


