---
title: "Homework 1 assignment - Data Visualization"
author: "CongW"
date: "September 16, 2014"
output: html_document
published: false
tags: hw1
---

##Summary
#Data Set:  NFL regular games 2014.
This data set contains data of 8174 games in 2014, displaying records on Date, Visitor team name, Visitor Score, Home team name, Home score, Line and Total Line

##Basic statistics
The following tables shows how many game a team played as a visitor team during 2014.
```{r,echo = FALSE,cache = TRUE}
library(ggplot2)
regular = read.csv("regular.csv")
summary(regular$Visitor)
```
The following tables shows how many game a team played as a home team during 2014.
```{r,echo = FALSE,cache = TRUE}
regular = read.csv("regular.csv")
summary(regular$Home.Team)
```

A basic statistics of visitor score.
```{r, echo=FALSE}
summary(regular$Visitor.Score)
```

A basic statistics of home score.
```{r, echo=FALSE}
summary(regular$Home.Score)
```

Following histomgram shows how different are the scores of home team and visitor team.
```{r, echo=FALSE}
df <- data.frame(cond = factor( rep(c("Visitor","Home"), each=8174) ),score = c(regular$Visitor.Score,regular$Home.Score)) 
ggplot(df, aes(x=score, fill=cond)) + geom_histogram(binwidth=3, alpha=.5, position="identity")
```

The plots above showing a difference between the home team scores and the visitor team scores. 