---
title: "Homework 1 assignment-JanetWenjuanDeng"
author: wd2234
output: html_document
pulished: false 
tags: hw1
---

This famous (Fisher's or Anderson's) iris data set consists of 50 samples from each of three species of flower(Setosa, Virginica and Vericolor). There are four variables that measure four features of each flower sample: petal length, petal width, septal length, and septal width, in centimeters.Since it is a R dataset package, there is no need to import data. 

```{r}
attach(iris)
library(ggplot2)
```

#scatterpoint plot
There is a positive and approximately linear association between sepal length and width. Meanwhile, Virginica has a relatively higher sepal length at the same level of sepal width, compared to versicolor and setosa.

```{r, echo=FALSE}

ggplot(data=iris,aes(x=Sepal.Width,y=Sepal.Length,colour=Species))+ geom_point(aes(shape=Species),size=2.5) + ggtitle("Sepal Length and Width")+stat_smooth(method="lm",formula=y~x,se=FALSE)

```

There is a positive and approximately linear association between petal length and width. Meanwhile, Virginica has the largest size petal, with highest level of both petal length and width on average. Versicolor comes second place, and setosa tends to have very small-sized petal. 

```{r, echo=FALSE}

ggplot(data=iris,aes(x=Petal.Width,y=Petal.Length))+ geom_point(aes(color=Species,shape=Species),size=3.5) + ggtitle("Petal Length and Width")+stat_smooth(method="lm",formula=y~x,se=FALSE,color="grey",size=1)
                                                                                                                                                  
```

#Boxplot,jitter plot and violin plot
These plots also shows that virginica has a relatively high sepal length, versicolor comes second, and setosa comes last.
```{r,echo=FALSE}
ggplot(data=iris,aes(x=Species,y=Sepal.Length)) + geom_boxplot(aes(fill=Species))+geom_jitter(aes(colour=Species),size=3)+ggtitle("Distribution of Sepal Lengths in box plot")

ggplot(data=iris,aes(x=Species,y=Sepal.Length)) +geom_jitter(aes(colour=Species),size=3)+geom_violin(aes(fill=Species))+ggtitle("Distribution of Sepal Lengths in violin plot")

````

#Bar plot
According to the graph, the sepal length follows a normal distribution. Viginica has a higher variance of sepal length. The sepal length of vesicolor and setosa are less volatile. The mean for virginica is about 6.3, for versicolor is about 6, and for setosa is about 5. 

```{r,echo=FALSE}
p<-ggplot(data=iris,aes(x=Sepal.Length)) + geom_histogram(aes(fill=Species))+facet_wrap(~Species) + ggtitle("Distribution of Sepal Lengths")
suppressMessages(print(p))
```
