Title: Ben Lee HW1
published: false
tags: hw1
========================================================
## Dataset: Daily Temperatures from 1995-2013

Read in the data
```{r}
data <- read.table("C:/Users/blee/Documents/QMSS Fall 2014/DataViz/qmssviz/Lab1/Daily_Temperature_1995-2013_DELHI.txt")
```

Add headers for the columns
```{r}
colnames(data)<-c("Month", "Day","Year","Temperature")
```

Plot of Daily Temperatures By Month
Cleaned data for plotting. This includes converting date information to the proper DATE format. 
```{r}
data$Time <- paste(data$Year, data$Month, data$Day, sep="-")
data$Date<-as.Date(data$Time)
```
Added a color scheme for month
```{r}
colorsch<-data.frame(c(1:12),c(rep(1,3),rep(2,3),rep(3,3),rep(4,3)))
```

## Scatterplot which displays the cyclical nature of temperature
```{r}
plot(data$Date[data$Temperature!=-99], data$Temperature[data$Temperature!=-99], main="Temperature by Day for 1995-2013",xlab="Date", ylab="Temperature (Fahrenheit)", ylim = c(min(data$Temperature[data$Temperature!=-99]),max(data$Temperature[data$Temperature!=-99]) ), type ="p", pch=20, col = colorsch[data$Month,2])

legend("bottom", legend=c("Jan/Feb/Mar","Apr/May/Jun","Jul/Aug/Sep","Oct/Nov/Dec"), fill=c(1,2,3,4),cex=0.85,horiz=TRUE, bty="n", x.intersp)
```

Scatterplot of average yearly temperature with simple linear regression line
Data Preparation. Removed data from 2013 because it is incomplete. 
```{r}
avtemp<-tapply(data$Temperature,data$Year, mean)
avtemp<-avtemp[1:18]
tempavg<-data.frame(as.numeric(as.character(names(avtemp))),avtemp)
colnames(tempavg)<-c("Year","AverageTemperature")
```

Table of average temperatures by year
```{r}
avtemp
```


## Scatterplot which displays average temperature by year and fits a linear regression line
The slope is not statistically significant at even the 10% significance level. This tells us that we cannot reject the null hypothesis that there is no relation between year and temperature. In other words, there is no strong evidence to support a claim that temperature and year are dependent (on average). 
```{r}
plot(tempavg[,1],tempavg[,2], main="Average Temperature by Year", 
xlab="Year", ylab="Temperature (Fahrenheit)", pch=20)
regmod<-lm(tempavg$AverageTemperature~tempavg$Year)
abline(regmod, col="red")
text(2007, 75.5, "Regression Line:\n y = 75.39678 +0.03855x \n p-value<0.5",
     col="red",cex = .8)
```

Regression Line details
```{r}
summary(regmod)
```
