---
title: "Homework 1 assignment"
author: yb2334
output: html_document
published: false
tags: hw1
---

# Data Visualization Homework 1
## Yannan Bu (yb2334)
========================================================

### Dataset: 
    responses-2014-09-05.csv

### Description: 
    This dataset contains information about students of QMSS Data Visualization course collected in the first session on Sept 5, 2014 (http://bit.ly/qmssvizt0). Most of the questions concern about students' basic information and software experience. 

### 1. Preparation: Getting and Cleaning Data

1.1 Read data and have a preliminary sense of the general picture. 
```{r}
setwd("E:/QMSS/Everything abt R/qmssviz/labs/questionnaire")
data <- read.csv("responses-2014-09-05.csv", header = T, stringsAsFactors = F)
str(data)
head(data, 5)
```

1.2 Clena data
Change classes of some variables and variable names. 
```{r}
data$Program <- as.factor(data$Program)
colnames(data)[4] <- "Twitter"
data$Twitter <- as.factor(data$Twitter)
```

Tranform multiple choice variable "Baseline.experience" into dichotomous variables. 
```{r}
data <- as.data.frame(sapply(data, gsub, pattern = "cvs, svn, hg, bzr", replacement = "cvs; svn; hg; bzr"))
data <- as.data.frame(sapply(data, gsub, pattern = "ec2, s3", replacement = "ec2; s3"))
data <- as.data.frame(sapply(data, gsub, pattern = "html, css, js", replacement = "html; css; js"))
lev <- levels(factor(data$Baseline.experience))
lev <- unique(unlist(strsplit(lev, ", ")))
name <- gsub(" ", "_", paste("Base.ex", lev, sep = "."))
result <- matrix(data = "0", nrow = length(data$Baeline.experience), ncol = length(lev))
char <- as.character(data$Baseline.experience)
for (i in 1:length(lev)) {
  result[grep(lev[i], char, fixed = T), i] <- "1"
}
result <- data.frame(result, stringsAsFactors = T)
colnames(result) <- name
data <- cbind(data, result)
```

### 2. Exploratory Analysis

All the analysis are done with the "ggplot2 package.
```{r}
library(ggplot2)
```


