---
title: "Homework 1 assignment"
author: yb2334
output: html_document
published: false
tags: hw1
---

# Data Visualization Homework 1

## Yannan Bu (yb2334)
========================================================

### Dataset: 
    responses-2014-09-05.csv

### Description: 
    This dataset contains information about students of QMSS Data Visualization course collected in the first session on Sept 5, 2014 (http://bit.ly/qmssvizt0). Most of the questions concern about students' basic information and software experience. 

### 1. Preparation: Getting and Cleaning Data

1.1 Read data and have a sense of the general picture. 
```{r}
setwd("E:/QMSS/Everything abt R/qmssviz/labs/questionnaire")
data <- read.csv("responses-2014-09-05.csv", header = T, stringsAsFactors = F)
str(data)
```

1.2 Clena data

```{r}
#Change classe of variable "Program" from character to factor. 
data$Program <- as.factor(data$Program)

table(data$Program)

#Combine "EdD" into the catogory of "Other PhD", "Statistics" into "Stat", and QMSS third semester students to the group "QMSS Second semester+"

data <- as.data.frame(sapply(data, gsub, pattern = "EdD", replacement = "Other PhD"))
data <- as.data.frame(sapply(data, gsub, pattern = "Statistics", replacement = "Stat"))
data <- as.data.frame(sapply(data, gsub, pattern = "QMSS third semester", replacement = "QMSS Second semester +"))
data <- as.data.frame(sapply(data, gsub, pattern = "QMSS Third semester", replacement = "QMSS Second semester +"))


#Tranform multiple choice variable "Baseline.experience" into dichotomous variables.

data <- as.data.frame(sapply(data, gsub, pattern = "cvs, svn, hg, bzr", replacement = "cvs; svn; hg; bzr"))
data <- as.data.frame(sapply(data, gsub, pattern = "ec2, s3", replacement = "ec2; s3"))
data <- as.data.frame(sapply(data, gsub, pattern = "html, css, js", replacement = "html; css; js"))
lev <- levels(factor(data$Baseline.experience))
lev <- unique(unlist(strsplit(lev, ", ")))
name <- gsub(" ", "_", paste(lev, sep = "."))
result <- matrix(data = 0, nrow = length(data$Baseline.experience), ncol = length(lev))
char <- as.character(data$Baseline.experience)
for (i in 1:length(lev)) {
  result[grep(lev[i], char, fixed = T), i] <- 1
}
result <- data.frame(result, stringsAsFactors = T)
colnames(result) <- name
data <- cbind(data, result)


#Create a new variable named "TotalBE" indicating the total number of baseline experience softwares or languages.

data$TotalBE <- rowSums(data[14:45])

#Create a new data frame to sum up the number of students that have previous experience with each software or language.

tools <- as.data.frame(sapply(data[, 14:45], function(v) {
  sel <- as.numeric(v==1)
  sum(sel)
}))
colnames(tools) <- "count"

#And sort the data frame by descending order
tools <- tools[ order(-tools[, 1]), , drop = F]
tools$name <- rownames(tools)

```

### 2. Exploratory Analysis

Note: All the analysis are done with the "ggplot2" package.
```{r, warning = FALSE}
library(ggplot2)
```


2.1 Student Background Summary
The first plot shows a histogram of the number of students from each program.
```{r, fig.align='center'}
plot1 <- qplot(Program, data = data, color = Program, size = I(1.5), main = "Student Background Summary", ylab = " ")
last_plot() + theme(axis.text.x = element_blank(), title = element_text(size = 12)) + coord_fixed(ratio = 0.3)
```

2.2 Baseline Experience Analysis

The second plot demonstrates top 10 most popular computer tools among Data Visualization students.
```{r}
sub_tool <- tools[10:1, ]
plot2 <- ggplot(sub_tool, aes(x= factor(name, levels=unique(name)), y = count)) +
  geom_bar(stat="identity", alpha = 1, position = "identity") +
  coord_flip() +
  xlab(" ") +
  ylab(" ") +
  ggtitle("Top 10 Most Popular Computer Tools") +
  guides(fill = FALSE)
print(plot2)
```

The third plot displays number of tools used before of each student, grouped by program.
```{r, fig.align='center'}
plot3 <- qplot(Program, TotalBE, data = data, color = Program, geom = "point", ylab = "Total Number of Baseline Experience Tools", main = "Baseline Experience Summary")
last_plot() + theme(axis.text.x = element_blank()) + coord_fixed(ratio = 0.25)
```

2.3 Software Experience Analysis

The fourth plot shows distribution of level of students' experience with R graphics.
```{r, fig.height=6,fig.width=6, fig.align='center'}
plot4 <- qplot(data[, 6], data = data, xlab = "Level of Experience with R Graphics", ylab = " ", main = "Distribution of R Graphics Experience") +
  facet_wrap(~Program, ncol = 3)
print(plot4)
#last_plot() + coord_fixed(ratio = 0.1)
```

The fifth plot is a pie chart of students' version control experience.
```{r, fig.align='center'}
plot5 <- ggplot(data, aes(x = factor(1), fill = factor(data[, 9]))) + 
  geom_bar(width = 1) +
  xlab(" ") +
  ylab(" ") +
  ggtitle("Proportional Distribution of Version Control Experience") +
  scale_fill_discrete("Level of Experience")
plot5 + coord_polar(theta = "y")
