---
title: "Homework 1 assignment"
author: "sl3641"
date: "September 16, 2014"
output: html_document
published: false
tags: hw1
---

CNN had an article this month about the World Health Organization's report on suicides in different countries:  
http://www.cnn.com/2014/09/05/world/who-global-suicide-report/

According to the article, South Korea had one of the highest suicide rates (after Guyana and North Korea).   
Among OECD countries, South Korea has had the highest reported suicide rates in the last 10 years.

I decided to look more closely at the data provided by WHO and combine it with data on other social and economic factors.

<br>
```{r, echo=FALSE}
suicideRate <- read.csv("/Users/User/Documents/qmssviz_other/Suicide Data/suicide_gdp.csv")
suicideRate$logGDPpc <- log(suicideRate$GDPperCap)
```        

```{r}
head(suicideRate)
str(suicideRate)
```

<br>

Below we have a scatterplot of 2010 suicide rates for 59 countries and GDP per capita.

```{r, echo=FALSE}
require(ggplot2)

ggplot(suicideRate, aes(x=logGDPpc, y=SuicideRate)) +
  geom_point(shape=1) +
  ggtitle("2010 Suicide Rates vs. GDP per Capita") +
  xlab("GDP per Capita, log scale") +
  ylab("Suicide rates (per 100,000)") +
  geom_text(aes(label=Country), hjust=0, vjust=0, size=3) +
  annotate("rect", xmin = 9.8, xmax = 10.7, ymin = 25, ymax = 27, alpha = .2, fill="yellow")
```

<br>
The plot below shows an increase in suicide rates in South Korea over time (since 1990).
```{r, echo=FALSE}
kor <- read.csv("/Users/User/Documents/qmssviz_other/Suicide Data/southkorea.csv")

require(ggplot2)

ggplot(kor, aes(x=year, y=suicideRate)) +
  geom_line() +
  geom_point() +
  ggtitle("South Korea Suicide Rates 1990 - 2011") +
  xlab("Year") +
  ylab("Suicide rates (per 100,000)")
```   

However, overall mortality rates have declined during this period.

```{r, echo=FALSE}
ggplot(kor, aes(x=year, y=mortality)) +
  geom_line() +
  geom_point() +
  ggtitle("South Korea Mortality Rates 1990 - 2011") +
  xlab("Year") +
  ylab("Mortality rates (per 100,000)")
```   



Data sources:

http://apps.who.int/healthinfo/statistics/mortality/whodpms/

http://data.worldbank.org/indicator/NY.GDP.PCAP.CD
