---
title: "hw1"
author: "David Tomczyk"
date: "September 18 2014"
output: html_document
published: FALSE
tags: hw1
---
```{r init, include=FALSE}

library(ggplot2)

setwd("~/qmssviz/Labs/questionnaire")
responses <- read.csv("responses-2014-09-05.csv",stringsAsFactors=F)
```
---

My first foray into R and ggplot2

---

# Who is in the class?

I thought it would be useful to take a look at who is taking this class. Its a QMSS class, but how many non-QMSS students are here?


```{r ggplot, fig.width=12, fig.height=3, message=FALSE}

program2.df <-as.data.frame.table(table(responses[[2]]),stringsAsFactors=FALSE)
program2.df <- program2.df[-8,]
program2.df[8,"Freq"] <- 2
program2.df <- program2.df[-6,]
program2.df[6,"Freq"] <- 2
names(program2.df)[1] <- "Program"
names(program2.df)[2] <- "Count"
program2.df[4,1] <- "QMSS 1st sem."
program2.df[5,"Program"] <- "QMSS 2nd Sem.+"
program2.df[6,"Program"] <- "QMSS 3rd Sem."

ggplot(data=program2.df, aes(x=Program, y=Count, fill=Program)) + geom_bar(stat="identity")+guides(fill=FALSE)+ggtitle("Class Make Up")

```
---

**What I like:**  
-I was able to manipulate the dataset in r to get counts(not necessary but useful), condense similar categories, change axis/title

**What I don't like:**  
-The colors  
-I think I'd like to order the data, though I'm not sure how  
-I'd like to put count number on top of the bars

---

#Twitter users among the class, by program

Which program/cohort has the highest proportion of twitter users?

```{r ggplot2, fig.width=16, fig.height=8, message=FALSE}

progtwit <- responses[,c(2,4)]
progtwit[14,"Program"] <- "Stat"
progtwit[28,"Program"] <- "QMSS Third semester"
progtwit2.df <-as.data.frame.table(table(progtwit),stringsAsFactors=FALSE)
progtwit2.df[4,"Program"] <- "QMSS 1st Sem."
progtwit2.df[5,"Program"] <- "QMSS 2nd Sem.+"
progtwit2.df[6,"Program"] <- "QMSS 3rd Sem."
progtwit2.df[12,"Program"] <- "QMSS 1st Sem."
progtwit2.df[13,"Program"] <- "QMSS 2nd Sem.+"
progtwit2.df[14,"Program"] <- "QMSS 3rd Sem."

ggplot(progtwit, aes(Program, fill=Do.you.use.twitter.)) + geom_bar() + ggtitle("Twitter Users by Program")

```
---
#look into Rcharts, have a button to switch between yes/no's in each category. Also think about adding proportions and/or frequency to the top of the bars. USE TWITTER BLUE(for users)/GREY(for not)! Consider: Dot plot, not bar chart

**What I like:**  
-The colors, I guess  
-More interesting than a simple bar chart

**What I don't like:**  
-Still a pretty simple bar chart  
-The sizes of the legend/axis labels/title  
-Might be useful to see specific count within bar groups, though that might look cluttered  

---

#People who know how to use version control

I had a lot of trouble with git over the week, so I wanted to look at who the other git noobies were.

``` {r ggplot3, fig.width=4, fig.height=7, message=FALSE}

vcs <-factor(responses[[9]])
vcs.table <-table(vcs)
vcs.df <- as.data.frame.table(vcs.table)

ggplot(data=vcs.df, aes(x=vcs, y=Freq), main="Version Control Experience") + geom_bar(stat="identity") + ggtitle("Experience With Version Control")

```
----
Looks kind of bad...what if we use a more appropriate scale?

#People who know how to use version control

``` {r ggplot4, fig.width=4, fig.height=3, message=FALSE}

ggplot(data=vcs.df, aes(x=vcs, y=Freq)) + geom_bar(stat="identity") + ggtitle("Experience With Version Control")

```