---
title: "hw1"
author: "David Tomczyk"
date: "September 18 2014"
output: html_document
published: FALSE
tags: hw1
---
```{r init, include=FALSE}

library(ggplot2)

setwd("~/qmssviz/Labs/questionnaire")
responses <- read.csv("responses-2014-09-05.csv",stringsAsFactors=F)
```
---

My first foray into R and ggplot2

---

# Who is in the class?

I thought it would be useful to take a look at who is taking this class. Its a QMSS class, but how many non-QMSS students are here?


```{r ggplot, fig.width=12, fig.height=3, message=FALSE}

program2.df <-as.data.frame.table(table(factor(responses[[2]])))
program2.df <- program2.df[-8,]
program2.df[8,"Freq"] <- 2
program2.df <- program2.df[-6,]
program2.df[6,"Freq"] <- 2

ggplot(data=program2.df, aes(x=Var1, y=Freq), main="Class Make Up") + geom_bar(stat="identity")

#Rename QMSS's so they fit better. Change Var1 column name to Program

#qplot(x=ProgramType, data=program, geom="bar", fill=ProgramType, main="Class Make Up")

```
---

What I like:  
-Its my first graph in R/ggplot2

What I don't like:  
-The colors  
-I wish the x axis labels were more spread out, or just gone  
-The x axis titles should simply be "Program"  
-I think I'd like to order the data, though I'm not sure how  
-Two redundant categories I would like to merge together, but didn't have time to learn how
-The colors

#People who know how to use version control

I had a lot of trouble with git over the week, so I wanted to look at who the other git noobies were.

``` {r ggplot2, fig.width=3, fig.height=7, message=FALSE}

vcs <-factor(responses[[9]])
vcs.table <-table(vcs)
vcs.df <- as.data.frame.table(vcs.table)

ggplot(data=vcs.df, aes(x=vcs, y=Freq), main="Version Control Experience") + geom_bar(stat="identity")
```
----
Looks kind of bad...what if we use a more appropriate scale?

#People who know how to use version control

``` {r ggplot3, fig.width=3, fig.height=3, message=FALSE}

ggplot(data=vcs.df, aes(x=vcs, y=Freq)) + geom_bar(stat="identity")

```