---
title: "Homework 1 Assignment"
author: "yw2668"
date: "September 19, 2014"
output: html_document
published: false
---

# I am a R beginner. I intend to make a picture showing different levels of software experience in [R, data manipulation and modeling] of people in different programs. ##

##Here I go! re-knitting # 

```{r}
setwd("/Users/Emily/Documents/Columbia/Data Visualization")
r<-read.csv("responses-2014-09-05.csv",stringsAsFactors=F)

```

```{r}

str(r)
```
<br/>
##1st try##
```{r}
library(ggplot2)
ggplot(r,aes(x=Program, y=Software.experience..the.smaller.list...R..data.manipulation.and.modeling.))+geom_point(shape=1)
```

##TERRIBLE##
 
 <br/>
```{r}
###give different levels numbers###
r$soft.r[r$Software.experience..the.smaller.list...R..data.manipulation.and.modeling.=="Expert"]<-3
r$soft.r[r$Software.experience..the.smaller.list...R..data.manipulation.and.modeling.=="Confident"]<-2
r$soft.r[r$Software.experience..the.smaller.list...R..data.manipulation.and.modeling.=="A little"]<-1
r$soft.r[r$Software.experience..the.smaller.list...R..data.manipulation.and.modeling.=="None"]<-0
table(r$soft.r)
```
<br/>
##2nd try##
```{r}
r$soft.r.level<-ordered(r$soft.r, level=c(0,1,2,3),labels=c("None","A little", "Confident", "Expert"))
ggplot(r,aes(x=Program, y=soft.r.level))+geom_point(shape=1)

```

##Now the labels of y-axis are listed in order##
##But it is still not what I want##
<br/>
#3rd try##
```{r}
##trying to count numbers, but it seems that I used wrong function##

r$sum.soft.r<-sum(r$soft.r)
ggplot(r,aes(x=soft.r.level, y=sum.soft.r, fill=Program))+geom_bar(stat="identity", position="dodge")
```

##Close to what I want, but the counts of levels hasn't been represented because I used wrong function##
<br/>
#4th try#
```{r}
qplot(x=soft.r.level, fill=Program,data=r,geom="bar",stat="bin")
ggplot(r,aes(x=soft.r.level, fill=Program))+geom_bar(stat="bin", position="dodge")+xlab("Levels in R, Data Manipulation & Modeling") + 
    ylab("Count") + 
    ggtitle("Different Levels in R of Different Programs")
```


##Remaining Questions:
1. Why the pictures are different using qplot and ggplot?

2.How to add count numbers over each bar?
