
title: "revised hw1"
author: "Lanxiang Li"
date: "September 24, 2014"
output: html_document
published:false
tags:hw1

## revised hw1
### Lanxiang_Li

```{r}
setwd("~/Desktop/qmssviz/Lab1")
dd <- read.csv("Donation-Disease.csv")
```
####1 Death of different diseases--scatter diagram
```{r}
library(ggplot2)
qplot(x = Name, y = Death, data = dd, color = Name)
```

####2 MoneyRaised of different diseases--scatter diagram
```{r}
library(ggplot2)
qplot(x = Name, y = MoneyRaised, data = dd, color = Name)
```

####3 MoneyRaised & Death--scatter diagram
```{r}
library(ggplot2)
plot(dd$Death,dd$MoneyRaised)
```

####4 MoneyRaised & Death--line graph
```{r}
library(ggplot2)
qplot(dd$Death, dd$MoneyRaised, geom="line")
```

####5 Death of diseases--bar graph
```{r}
library(ggplot2)
qplot(Name, Death, data=dd, geom="bar", stat="identity",color = Name)
```

####6 MoneyRaised of diseases--bar graph
```{r}
library(ggplot2)
qplot(Name, MoneyRaised, data=dd, geom="bar", stat="identity",color = Name)
```

#### my comment: I like the bar graph!Comparing to the scatter one, I think the bar graph is more clear and beautiful. It can show the data in such a beautiful way~

####7  Death of diseases--Box plot
```{r}
library(ggplot2)
plot(dd$Name, dd$Death)
qplot(dd$Name, dd$Death, geom="boxplot")
boxplot(Death ~ Name, data = dd)
```

####8 Death$MoneyRaised--linear 
```{r}
library(ggplot2)
ggplot(dd, aes(x=Death, y=MoneyRaised)) +
geom_point(shape=1)
```
```{r}
library(ggplot2)
ggplot(dd, aes(x=Death, y=MoneyRaised)) +
geom_point (shape=1) +
geom_smooth(method=lm)
```

####my comment: it seem that the death of a disease does not have much relevance with the MoneyRaised.

####9 MoneyRaised of Diseases--pie chart
```{r}
library(ggplot2)
par(mar=c(8,8,8,8),font=4)
pct <- round (dd$MoneyRaised/sum(dd$MoneyRaised)*100)
lbls<- paste(dd$Name, " ",pct, "%", sep="")
pie(dd$MoneyRaised, labels=lbls,col=rainbow(length(lbls)), main="pie chart of MoneyRaised")
```

####10 Death of Diseases--pie chart
```{r}
library(ggplot2)
par(mar=c(8,8,8,8),font=4)
pct <- round (dd$Death/sum(dd$Death)*100)
lbls<- paste(dd$Name, " ",pct, "%", sep="")
pie(dd$Death, labels=lbls,col=rainbow(length(lbls)), main="pie chart of Death")
```

