---
Title: "Homework 1 Assignment"
author: Andy Tang, awt2118
output: html_document
published: false
tags: hw1
========================================================
```{r}
setwd ("/Users/awtang/Desktop/qmssviz/qmssviz")
measures<-read.csv("bodyfat.csv", stringsAsFactors=F)
summary(measures)
```

```{r}
library(ggplot2)
```

In this scatter plot, we see the trend of obese people having highest abdomen size and neck size, followed by overweight people, healthy people, and underfat people. 
```{r fig.width=10, fig.height=8 }
g<-ggplot(measures, aes(Abdomen, Neck,))

g+geom_point(aes(color=FatStat), size=4, alpha=1/2)+labs(title="Neck and Abdomen Size by Fat Status")+labs(x=("Abdomen size in inches"), y="Neck size in inches")
```

When looking at this scatter plot, we see that age is pretty even distributed. 
```{r fig.width=10, fig.height=8}
g+geom_point(aes(color=Age), size=4, alpha=1/2)+labs(title="Neck and Abdomen Size by Age Groups")+labs(x=("Abdomen size in inches"), y="Neck size in inches")
```


These facets show the distributions of neck size and abdomen size in the 4 different fat status categories and the 4 age groups. 
```{r fig.width=10, fig.height=8}
g+geom_point()+facet_grid(~FatStat)+labs(title="Neck and Abdomen Size By Fat Status")+labs(x=("Abdomen size in inches"), y="Neck size in inches")
g+geom_point()+facet_grid(~Age)+labs(title="Neck and Abdomen Size By Age Groups")+labs(x=("Abdomen size in inches"), y="Neck size in inches")
qplot(FatStat, data=measures, fill=Age)+labs(title="Age Groups Across Fat Status Groups")+labs(x=("Fat Status"), y="Count")
```
