---
title: "hw1"
author: "John Peng"
output: html_document
published: false
tags: hw1
---


This dataset(firms) comprises a set of companies and looks at their industry(industry.recode), employee size(empsize), and assets(assets).

**Plotting # of employees x assets**

```{r, echo=FALSE, warning=FALSE}
setwd('C:/Users/John/Desktop/Data Visualization QMSS/qmssviz')
library(ggplot2)
load('soc504_sp13_pager.RData')

#plotting # of employees x assets
ggplot(firms, aes(x=empsize, y=assets)) +
  geom_point(shape=1, size =1) +
  xlim(0,1000) +
  ylim(0,150)
```

**Hiding outlier by adjusting plot limits**

```{r, echo=FALSE, warning=FALSE}
ggplot(firms, aes(x=empsize, y=assets)) +
  geom_point(shape=1, size =1) +
  xlim(0,200) +
  ylim(0,80) +
  xlab("# of Employees (thousands)") +
  ylab("Assets (millions of dollars)")
```

**Zooming in on clustered data points at lower values**

```{r, echo=FALSE, warning=FALSE}
ggplot(firms, aes(x=empsize, y=assets)) +
  geom_point(shape=1, size =1) +
  xlim(0,40) +
  ylim(0,20) +
  xlab("# of Employees (thousands)") +
  ylab("Assets (millions of dollars)")
```

- Number of employees does not seem to be a very accurate indicator of assets, but it does seem to be an indicator of the maximum assets for a company of that size, since the highest asset company is increasing as number of employees increasing. This is possibly due to the lack of success of a high proportion of companies across all employee sizes. However, a more detailed analysis would be needed to determine if there is or isn't an association for these variables.


**Barplot of industry breakdown for businesses**

```{r, echo=FALSE, warning=FALSE}
ggplot(firms, aes(x=industry.recode)) + 
  geom_bar(aes(fill=industry.recode)) +
  scale_fill_hue(name="Industry", l=20) +
  xlab("Industry") +
  ylab("# of Companies") +
  theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())
```

- Most of the companies in the sample are in retail trade plus finance & services. It would be interesting to see in future analyses whether the high proportion of low-asset companies is due to this. Perhaps the assets of the businesses in some or one of these areas are especially volatile.


**Density plot of company assets distribution**

```{r, echo=FALSE, warning=FALSE}
ggplot(firms, aes(x=assets)) + 
  geom_density(alpha=.2, fill="#FF6666") +
  xlim(0,100) +
  ylim(0,0.4)
```

- A large proportion of companies in the sample have assets below $10 million.