---
title: "Homework 1, Rachel Smith"
author: rls2191
output: html_document
published:false
tags:hw1
---
I'm interested in investigating the dataset of health facilities in Niger. The dataset contains  information about the location, number of full-time doctors, nurses, and lab techs, and the management type for 44 health facilities. Some descriptives, including the number of facilities with the value NA for doctors and nurses:
```{r}
health_full <- read.csv("../../Lab1/sample_health_facilities.csv")
summary(health)
```
I'd like to graph the number of doctors vs. the number of nurses, but some entries have a value of NA for one or both of those pieces of information. Points with an NA value for the number of doctors or nurses are dropped, a total of 4 points.
```{r}
table(is.na(health_full$num_doctors_fulltime), is.na(health_full$num_nurses_fulltime))
health <- health_full[!(is.na(health_full$num_doctors_fulltime) | is.na(health_full$num_nurses_fulltime)), ]
```

Here is a scatter plot of doctors vs. nurses. The dot colors refer to the region of the clinic. 
```{r}
plot(health$num_doctors_fulltime, health$num_nurses_fulltime, main="Doctors and Nurses in Niger by Region", xlab="Number of Full Time Doctors", ylab="Number of Full Time Nurses", xlim=c(0,4), ylim=c(0,9), pch=19, col=health$zone)
legend(2.75, 8.75, unique(health$zone), pch=19, col=unique(health$zone))
```

Here is the same plot with the locations labeled:
```{r}
plot(health$num_doctors_fulltime, health$num_nurses_fulltime, main="Doctors and Nurses in Niger by Region", xlab="Number of Full Time Doctors", ylab="Number of Full Time Nurses", xlim=c(0,4), ylim=c(0,9), pch=19, col=health$zone)
legend(2.75, 8.75, unique(health$zone), pch=19, col=unique(health$zone))
text(health$num_doctors_fulltime, health$num_nurses_fulltime, labels=health$lga, cex= 0.7)
```

Showing the same information with ggplot
```{r}
library(ggplot2)
ggplot(health, aes(num_doctors_fulltime, num_nurses_fulltime, color=zone)) + 
  geom_point(shape=9, alpha=2/3, fill=5) + 
  scale_color_hue(l=50) + 
  labs(x="Number of Full Time Doctors", y="Number of Full Time Nurses") + 
  theme(                              
      plot.background = element_blank(), 
      panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank(), 
      panel.border = element_blank(), 
      panel.background = element_blank(),
      axis.line = element_line(size=.4)
  )
```

