---
title: "hw1 huzhongjun"
author: "Zhongjun Hu"
date: "2014年9月17日"
output: html_document
---
# HW1
## HU
```{r}
library(ggplot2)
```
### Donation-Disease:
```{r,fig.height=3, fig.width=9}
setwd("~/Desktop/ZhongjunHuqmssviz/qmssviz/Lab1")
d <-read.csv("Donation-Disease.csv")
d$Death<-as.numeric(d$Death)
d$Death<-d$Death/1000
qplot(data=d,x = MoneyRaised,y=Death,color=Name)

ggplot
```

### UN population
(I noticed a very nice pic from one of our classmates. He taught me and I came back home trying by myself but still failed.His plot is way way way better than mine. No ideas why)
```{r, fig.height=7, fig.width=15}
library(RColorBrewer)

library(xlsx)
color <- brewer.pal(5, "Reds")
setwd("~/Desktop/ZhongjunHuqmssviz/qmssviz/Lab1")
un2<-read.xlsx("UN3.xlsx",sheetName="sheet1")
un2$Population_Thousand<-cut(as.numeric(un2$X2010),breaks=c(0,2000,5000,10000,20000,Inf),labels=c("<2000","2000~5000","5000~10000","10000~15000",">15000"))
qplot(data=un2,x=Longitude,y=Latitude,color=Population_Thousand) +geom_point(size=3) +scale_color_manual(values=color)

```
### Ages of Marriage
```{r}
setwd("~/Desktop/ZhongjunHuqmssviz/qmssviz/Lab1")
cb<-read.csv("GSS.2006.csv")
cb$Ages<-as.numeric(cb$agewed)
qplot(x=Ages, data=cb,geom="density",main="Distribution of How old were you when first married")
qplot(x=Ages, data=cb,geom="histogram",main="Distribution of How old were you when first married")
