---
title: "R Basics"
author: "Yang Yang"
---


Brief intro of RStudio, an integrated development environment (IDE)

Anything following a # is a comment

```{r}
# This is a comment, which is not executed.
# You can take notes this way, e.g.Commom dataset structure:each variable is a column, each observation is a row, and each type of observational unit is a table. 
```

R as a calculator
-------------------------

#### Basic Math
```{r}
1 + 1
4 / 2 * 3
```


#### R follows the order of operations PEMDAS Parenthesis Exponents Multiplication Division Addition Subtraction
```{r}
4 / (2*3)
```

#### R has built in math functions 
```{r}
sqrt(4)
```



Assign a variable
-------------------------
```{r}
a <- 1
b <- 2   ## you can also try b=2, which is another way of assigning a variable, but arrow is generally preferred
c <- a + b 
```

```{r}
c
```


#### R is case sensitive
```{r}
A <- 10
```

```{r}
A
a    ##remember, R is case sensitive! 
```

#### Remove objects 
```{r}
ls()  # R stores both data and data analysis output as objects. ls() returns a list of all objects in the current session
rm(a) 
```


Vectors
----------------------------------------

#### Create vectors *Note: R is case sensitive*
```{r}
A <- c(1,2,3,4,5)  ##Overwrite an object by assigning it something new
A
B <- 1:5; B #Commands are separated either by a ; or by a newline
C <- c(1, 2, 4, 3, 5); C 
```


#### Logical (hints for vector operations)
```{r}
A < 2

##check length
length(A); length(B)

A == B
A == C
A != 5
```


#### Vector Operations
```{r}
# add 1 to every element
A + 1

# multiplty 2 to every element
A * 2

# raise every element to second power
A ^ 2
```


#### More on Vector Operations (for vectors have the same length)
```{r}
# +, -, *, /
A+B

A/B 

```


```{r}
##get a specific element from vector A
A[3]

##remove a specific element from vector A
A[-1]

D<-A[-1]
D
length(D)
```


Introduction to data types and data.frame 
----------------------------------------

#### Set working directory and read dataset
It should be your own working directory in "..."; if you are using a R script, you should be fine to do as following two steps to read data: set working directory and read dataset

```{r}
#setwd("/Users/timz/Downloads")
demo <- read.csv("Donation-Disease.csv",
                 stringsAsFactors=F,
                # arg2=blah,
                # more args
                )
```

#### read.table and read.csv can read data stored as text files, delimited by nearly everything, e.g. comma delimited and tab delimited. 

```{r}
# give dataset a name as a R dataframe object 
demo<-read.csv("Donation-Disease.csv",stringsAsFactors=F)
```


```{r}
# class of demo
class(demo) # data frames have a matrix structure. Obs are arranged as rows and columns

# Look at the dataset 
head(demo)

# dimensions of demo
dim(demo)   #output: row column

# structure of demo, including data types for variables
str(demo)

# summary demo, returns distributional summaries of variables in the dataset
summary(demo)

# indexing rows, columns and cells in a dataframe
# object[row,column]
demo[1,1]

```

#### indexing rows, columns and cells in a dataframe. 
* object[row,column] 
* object$column.name
```{r}
# 
demo[1,1]
demo$Name
```
Install R Packages
----------------------------------------
#### To use packages in R, we must first install them using install.packages function, which downloads the packages from CRAN. 
You just need to install a package once for your RStudio, you only need to call this package in the future. 

```{r}
# install.packages("ggplot2") #package for elegant data visualization using the Grammar of Graphics
library(ggplot2)
?ggplot #Get R on-line help 

ggplot(demo, aes(x=MoneyRaised, y=Death)) +
    geom_point(shape=1) 

```

Brief intro R Markdown
-------------------------


Wrap-up:
-------------------------
 * Always remember to check working directory getwd() or set the right working directory setwd()
 * Use on-line help, get used to read tutorials on CRAN; also Google and Stack Overflow
 * Define meaningful&easy names for datasets and variables, remember R is case sensitive
 * Practice and create your own cheatsheet 
 * Quiz: To read dataset from other statistical analysis software, like Stata and SPSS, what packages shall we use? 

Useful Resources:
-------------------------
* R basic: http://cran.r-project.org/manuals.html
* R for Beginners
* R in a Nutshell
* Quick R: http://www.statmethods.net/
* Data Analysis Examples with R: http://www.ats.ucla.edu/stat/dae/



