---
title: "Visualizing Introspection by Country"
author: "rcquan"
date: "October 23, 2014"
output: pdf_document
---

# Objective

I am using the two most recent waves of the World Values Survey to examine the responses to the following survey question:

**How frequently do you think about meaning and purpose in life?**

The graph that I will be making will attempt to convey the following:

* A comparison of the response breakdown between countries within the same wave
* A comparison of the response breakdown within countries across different waves.

This information can be represented by a stacked horizontal bar plot that has the proportion of survey responses on the horizontal axis and the countries on the vertical axis.

Moreover, this particular survey question has five responses of interest:

1. Often
2. Sometimes
3. Rarely
4. Never
5. Don't know

Applying the appropriate color palette could help the viewer to more easily interpret the breakdown of each country's response. For example, we could use similar, positive colors for responses indicating a greater frequency of "thinking about meaning and purpose".

# Data Description

Before we wrangle our dataset into the appropriate long format for `ggplot`, let's first define the target shape of our data.

`{wave: "", country: "", meaning: "", proportion: ""}`

Here are possible values that each column could take:

* `wave`: "2005-2009", "2010-2014"
* `country`: "Afghanistan..."
* `meaning`: "Often", "Sometimes", "Rarely", "Never", "Don't Know"
* `proportion`: "numeric"

Since the objective is to compare all countries that have values in both waves, we do not yet know the exact length of this particular variable.

# Code

Let's declare our dependencies and define some helper functions so we can keep our environment clean:

```{r, message=FALSE}
library(ggplot2)
library(dplyr)
library(reshape2)
library(scales)
library(RColorBrewer)

########################
# FUNCTIONS
########################
selectVars <- function(df, code, name) {
    # Select survey variables to include in the subsetted data frame.
    # Args:
    #   df: a data.frame of the WVS survey data
    #   code: a vector of strings corresponding to survey codes
    #   name: a vector of strings corresponding to desired varname
    # Returns:
    #   A subsetted and named data.frame 
    if (length(code) != length(name)) {
    stop("Length of code and name vectors do not match.")
    }
    ## pre-allocate size of list
    varList <- vector("list", length(code))
    varList <- df[code]
    ## create data.frame from list
    dfNames <- data.frame(varList)
    names(dfNames) <- name
    return(dfNames)
}

factorizeVars <- function(df) {
    # Convert all variables into factors
    # Args:
    #   df: a data.frame
    # Returns:
    # a data.frame of variables of class "factor"
    data.frame(apply(df, 2, as.factor))
}

labelValues <- function(var, df, sep = ":") {
    # Reads in codebook from the current directory
    # and creates factor labels for the given variable.
    # If no codebook is found, no labeling is done.
    # Args:
    #   var: string. the name of the variable to be labeled
    #   df: data.frame. the data.frame containing the variable
    #   sep: string. the delimiter used in the codebook
    # Returns:
    #   A vector of class "factor" with the corresponding labels
    filename <- paste(var, ".txt", sep = "")
    if(file.exists(filename)) {
        dfLabels <- read.delim(filename, header = FALSE, sep = sep)
        labeledVar <- factor(df[, var],
                              levels = dfLabels$V1,
                              labels = dfLabels$V2)
        labeledVar
    } else {
        cat(sprintf("%s does not exist in the current directory. \n", filename))
        cat(sprintf("No labeling done for %s.", var))
        df[, var]
    }
}

labelDataFrame <- function(df, sep = ":") {
    # A wrapper for the above function to vectorize the labeling
    # process for all column variables.
    listLabeled <- sapply(colnames(df), function(var) {
        labelValues(var, df, sep = sep)
    })
    dfLabeled <- data.frame(listLabeled)
    return(dfLabeled)
}

reorderResponses <- function(df, var, newLevels) {
    # Reorder factor levels of response variable.
    # Args:
    #   df: data.frame. the data.frame containing the var to be reordered
    #   var: string. the name of variable to be ordered
    #   newLevels: a vector of strings to denote new var order
    # Returns:
    #   A data.frame
    df[, var] <- factor(df[, var],levels = newLevels)
    return(df)
}

handleMissingValues <- function(df, missingLabels) {
    # Remove missing responses from the dataset
    # Args:
    #   df: data.frame. the data.frame 
    #   missingLabels: string. a vector of reponse values to be removed
    # Returns:
    #   A data.frame with missing responsed removed
    df <- filter(df, !(meaning %in% missingLabels))
    return(df)
}

subsetByCountriesInBothWaves <- function(df, wave1, wave2) {
    # Subsets the data.frame by countries that have responses in both waves.
    # Args:
    #   df: data.frame. the data.frame 
    #   wave1: string. the year of the first wave to match against
    #   wave2: string. the year of the second wave to match against
    # Returns:
    #   A data.frame with countries that have responses in wave 1 and wave 2.
    firstWave <- df %>% filter(wave == wave1)
    secondWave <- df %>% filter(wave == wave2)
    
    firstWave <- firstWave[which(firstWave$country %in% secondWave$country), ]
    secondWave <- secondWave[which(secondWave$country %in% firstWave$country), ]
    
    df <- rbind(firstWave, secondWave)
    return(df)
}

reorderCountriesByProportion <- function(df, refWave) {
    # Reorders the levels of the country factor variable by
    # the proportion of responses from the reference wave.
    # Args:
    #   df: the data.frame
    #   refWave: string. the reference wave to sort by
    # Returns:
    #   A sorted data.frame
    countryIndex <- df %>%
        group_by(meaning) %>%
        filter(wave == refWave) %>%
        filter(meaning == "Often") %>%
        arrange(desc(proportion))
    
    countryIndex <- data.frame(countryIndex)
    df$country <- factor(df$country, levels = countryIndex$country)
    return(df)
}
```

# Visualization

Alright, now let's run the script to wrangle and visualize the dataset so we can answer our key question:

```{r, warnings=FALSE, message=FALSE}
## globals
newLevels <- c("Often", "Sometimes", "Rarely", "Never", "No answer",
               "Not asked in survey", "Missing: Unknown", "Don´t know")
missingLabels <- c("No answer", "Missing: Unknown", "Not asked in survey")

## reading in the data
df <- load("WVS_Longitudinal.rdata")
df <- `WVS_Longitudinal_1981-2014_rdata_v_2014_11_25`

df %>%
    # pre-processing
    selectVars(code = c("S002", "S003", "S020", "F001"),
               name = c("wave", "country", "year", "meaning")) %>%
    factorizeVars() %>%
    labelDataFrame() %>%
    reorderResponses(var = "meaning", newLevels = newLevels) %>%
    handleMissingValues(missingLabels = missingLabels) %>%
    
    ## get proportion of responses for each country by wave
    group_by(wave, country, meaning) %>%
        summarise(count = n()) %>%
        mutate(proportion = count/sum(count)) %>%
        arrange(meaning, desc(proportion)) %>%
    
    ## prep data for plotting
    subsetByCountriesInBothWaves(wave1 = "2005-2009",
                                 wave2 = "2010-2014") %>%
    reorderCountriesByProportion(refWave = "2005-2009") %>%
    
    ## generate the plot
    ggplot(aes(country, proportion, fill = meaning)) +
    geom_bar(stat = "identity", alpha = 7/8) + facet_grid(~wave) +
    ggtitle("How frequently do you think about meaning and purpose in life?") +
    xlab("Country") + ylab("Proportion of Survey Responses") +
    ## display proportion as percent
    scale_y_continuous(label = percent) +
    scale_fill_manual(values = rev(brewer.pal(n = 5, "RdYlBu"))) +
    coord_flip() + 
    ## minimalist theme
    theme_bw() +
    theme(legend.position = "bottom",
          legend.title = element_blank(),
          ## make reference lines darker
          panel.grid.major.x = element_line(color = "black"),
          panel.grid.major.y = element_blank(),
          panel.border = element_blank(),
          strip.background = element_blank(),
          axis.ticks = element_blank())
```
